---
title: "Chicago Pothole Complaints Analysis"
subtitle: "Identifying High-Priority Community Areas"
author: "Shaofeng Dai"
date: "2025/12/11"
format:
  html:
    toc: true
    toc-float: true
    theme: cosmo
execute:
  echo: false
  warning: false
---

## Introduction

This report visualizes the distribution of pothole service requests across Chicago community areas to help the Department of Transportation allocate repair resources more effectively.

## Data Source & Filtering

-   **Source**: Chicago Data Portal – 311 Service Requests\
-   **Filter**: `type_of_service_request` contains "Pothole"\
-   **Period**: Full available dataset (singel-year)\
-   **Records**: `r nrow(df)` requests after cleaning

## Data

```{r}
library(readr)
library(dplyr)
library(knitr)  
library(kableExtra)

processed <- read_csv("data/potholes.csv", show_col_types = FALSE)

sample_tbl <- processed %>%
  arrange(desc(creation_date)) %>%
  slice_head(n = 20)

sample_tbl %>%
  kable(format = "html", caption = "First 20 Rows of Cleaned Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Visualizations

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)

data <- read_csv("data/potholes.csv", show_col_types = FALSE) %>% 
  mutate(month = floor_date(creation_date, "month"))
```

```{r}
top20 <- count(data, community_area, sort = TRUE) %>% slice_head(n = 20)

ggplot(top20, aes(reorder(community_area, n), n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 20 Community Areas by Pothole Complaints",
       x = NULL, y = "Number of Requests") +
  theme_minimal(base_size = 14)
```

```{r}
top10 <- count(data, community_area, sort = TRUE) %>% slice_head(n = 10)

ggplot(top10, aes(x = "", y = n, fill = factor(community_area))) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  labs(title = "Top 10 Communities Share", fill = "Community Area") +
  theme_void()
```

```{r}
data <- data %>% mutate(week = isoweek(creation_date))   # 一年中的周 1-53

weekly <- count(data, week)

ggplot(weekly, aes(week, n)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(color = "darkred") +
  labs(title = "Pothole Requests by Week of Year",
       x = "ISO Week", y = "Requests") +
  theme_minimal()
```
